
\begin{thebibliography}{00}


% \bibitem{se1}Ali El Kamel, Hamdi Eltaief, Habib Youssef, On-the-fly (D)DoS attack mitigation in SDN using Deep Neural Network-based rate limiting, Computer Communications,Volume 182,2022, Pages 153-169, ISSN 0140-3664, https://doi.org/10.1016/j.comcom.2021.11.003.

% \bibitem{r3}El Kamel, A., Eltaief, H., Youssef, H. (2023). Improving the Routing Process in SDN Using a Combination of the Evidence Theory and ML. In: Abraham, A., Pllana, S., Casalino, G., Ma, K., Bajaj, A. (eds) Intelligent Systems Design and Applications. ISDA 2022. Lecture Notes in Networks and Systems, vol 717. Springer, Cham. https://doi.org/10.1007/978-3-031-35510-3\_26.

%########################### DDoS references
% ######### Definition
\bibitem{ddos2019} Iman Sharafaldin, Arash Habibi Lashkari, Saqib Hakak, and Ali A Ghorbani. Developing realistic distributed denial of service (ddos) attack dataset and taxonomy. In 2019 International Carnahan Conference on Security Technology (ICCST), pages 1–8. IEEE, 2019.

%######   Taxonomy

\bibitem{insdn} M.S. Elsayed, N.A. Le-Khac, and A.D. Jurcut, “InSDN: A novel SDN intrusion dataset,” IEEE Access, vol. 8, pp. 165263-165284, 2020.






% ######  Attacks types

\bibitem{ciciot2023} 
Neto, E.C.P.; Dadkhah, S.;
Ferreira, R.; Zohourian, A.; Lu, R.;
Ghorbani, A.A. CICIoT2023: A
Real-Time Dataset and Benchmark
for Large-Scale Attacks in IoT
Environment. Sensors 2023, 23, 5941.
https://doi.org/10.3390/
s23135941.
\bibitem{edgeiiot} 
MOHAMED AMINE FERRAG, OTHMANE FRIHA, DJALLEL HAMOUDA, LEANDROS MAGLARAS, AND HELGE JANICKE, Edge-IIoTset: A New Comprehensive Realistic
Cyber Security Dataset of IoT and IIoT
Applications for Centralized and
Federated Learning ,IEEE Access, 10.1109/ACCESS.2022.3165809.

% #### Effect
\bibitem{glass} 
Keerthiraj Nagaraj, Allen Starke, Janise McNair, GLASS: A Graph Learning Approach for Software Defined Network Based Smart Grid DDoS Security
\bibitem{analysis1} 
Nguyen Huu Thanh, Nguyen Ngoc Tuan, Dang Anh Khoa, Le Cong Tuan, Nguyen Trung Kien, Nguyen Xuan Dung, Ngo Quynh Thu, Florian Wamser, On Profling, Benchmarking and Behavioral Analysis of SDN 
Architecture Under DDoS Attacks, Journal of Network and Systems Management (2023), https://doi.org/10.1007/s10922-023-09732-5.
\bibitem{analysis2}  
Sukhveer Kaur,  Krishan Kumar, dNaveen Aggarwal, Analysis of DDoS Attacks in Software Defined Networking, 2022 IEEE Delhi Section Conference (DELCON)

\bibitem{stgcn} 
Yongyi Cao , Hao Jiang , Yuchuan Deng , Jing Wu , Pan Zhou , and Wei Luo, Detecting and Mitigating DDoS Attacks in SDN Using Spatial-Temporal Graph Convolutional Network, IEEE TRANSACTIONS ON DEPENDABLE AND SECURE COMPUTING, VOL. 19, NO. 6, NOVEMBER/DECEMBER 2022 .

% ######## detection

%################ DDoS detection and mitigation
% classical approaches
\bibitem{ali1} S. Shin and G. Gu, \emph{Attacking software-defined networks: A first feasibility study}, in ACM SIGCOMM Workshop on Hot Topics in
Software Defined Networking, 2013, pages:165–166.
\bibitem{ali2} R. Kandoi and M. Antikainen, \emph{Denial-of-service attacks in OpenFlow SDN networks}, in IFIP/IEEE International Symposium on Integrated
Network Management (IM), 2015, pages: 1322–1326.
\bibitem{ali3} L. Wei and C. Fung, \emph{FlowRanger: A request prioritizing algorithm for controller DoS attacks in software defined networks}, in IEEE
International Conference on Communications (ICC), 2015, pages:5254–5259.
\bibitem{ali4} 
R. Sahay, G. Blanc, Z. Zhang, and H. Debar, \emph{Towards autonomic DDoS mitigation using software defined networking}, in Network and
Distributed System Security (NDSS) Symposium, 2015.
\bibitem{ali5} 
N.-N. Dao, J. Park, M. Park, and S. Cho, \emph{A feasible method to combat against DDoS attack in SDN network}, in International Conference on
Information Networking (ICOIN), 2015, pages:309–311.
\bibitem{ali6} 
Lobna, Dridi and Zhani, Mohamed Faten, \emph{SDN-Guard: DoS Attacks Mitigation in SDN Networks}. 10.1109/CloudNet.2016.9, (2016).
%#########" ML based approaches
\bibitem{ml1} 
Ali, T.E.; Chong, Y.-W.; Manickam, S. Machine Learning Techniques to
Detect a DDoS Attack in SDN: A Systematic Review. Appl. Sci. 2023,
13, 3183. https://doi.org/ 10.3390/app13053183.
\bibitem{ml2} 
 Tonkal,  ̈O.; Polat, H.; Ba ̧saran, E.; C ̈omert, Z.; Kocao  ̆glu, R. Machine
Learning Approach Equipped with Neighbourhood Component Analysis
for DDoS Attack Detection in Software-Defined Networking. Electron-
ics 2021, 10, 1227. https://doi.org/10.3390/ electronics10111227.
\bibitem{ml3} 
 Makuvaza, Auther \& Jat, Dharm \& Gamundani, Attlee. (2021). Deep
Neural Network (DNN) Solution for Real-time Detection of Distributed
Denial of Service (DDoS) Attacks in Software Defined Networks
(SDNs). SN Computer Science. 2. 10.1007/s42979-021-00467-1.
\bibitem{ml4} 
 Ma, R.; Chen, X.; Zhai, R. A DDoS Attack Detection Method Based on
Natural Selection of Features and Models. Electronics 2023, 12, 1059.
https://doi.org/10.3390/ electronics12041059.
\bibitem{ml5} 
 Deepak Kumar, R.K. Pateriya, Rajeev Kumar Gupta, Vasudev Dehal-
war, Ashutosh Sharma,DDoS Detection using Deep Learning,Procedia
Computer Science,Volume.
\bibitem{ml6} 
M. S. Elsayed, N. -A. Le-Khac, S. Dev and A. D. Jurcut,
Machine-Learning Techniques for Detecting Attacks in SDN, 2019
IEEE 7th International Conference on Computer Science and Network Technology (ICCSNT), 2019, pp. 277-281, doi: 10.1109/ICCSNT47585.2019.8962519.
%#########" DL based approaches
\bibitem{dl1} 
 Tang, T.A. ; Mhamdi, L. ; McLernon, D. ; Zaidi, S.A.R. ; Ghogho, M. Deep Recurrent Neural Network for Intrusion Detection in
SDN-based Networks, In Proceedings of the 4th IEEE Conference on Network Softwarization and Workshops (NetSoft), Montreal, QC,
Canada, 25–29 June 2018 ; pp. 202–206.
\bibitem{dl2} 
 Dey, S.K. ; Md Rahman, M. Flow Based Anomaly Detection in Software Defined Networking : A Deep Learning Approach With
Feature Selection Method, In Proceedings of the 4th International Conference on Electrical Engineering and Information Communication
Technology (iCEEiCT), Dhaka, Bangladesh, 13–15 September 2018 ; pp. 630–635.
\bibitem{dl3} 
 Tang, T.A. ; Mhamdi, L. ; McLernon, D. ; Zaidi, S.A.R. ; Ghogho, M. (2016) Deep learning approach for Network Intrusion Detection
in Software Defined Networking, In Proceedings of the International Conference on Wireless Networks and Mobile Communications
(WINCOM), Fez, Morocco, 26–29 October 2016 ; pp. 258–263.
\bibitem{dl4} 

%#########" RL based approaches
\bibitem{dosRLReview} 
Hou, F., Sun, J., Yang, Q. et al. Deep reinforcement learning for optimal denial-of-service attacks scheduling. Sci. China Inf. Sci. 65, 162201 (2022). https://doi.org/10.1007/s11432-020-3027-0.
\bibitem{rl1} K. Malialis and D. Kudenko, Distributed response to network intrusions using multiagent reinforcement learning, Engineering
Applications of Artificial Intelligence, vol. 41, pp. 270–284, 2015.
\bibitem{rl2} Yandong Liu, Mianxiong Dong,Kaoru Ota, Jianhua Li, Jun Wu, Deep Reinforcement Learning based Smart Mitigation of DDoS Flooding
in Software-Defined Networks, IEEE 23rd International Workshop (CAMAD), 2018.
\bibitem{rl3} Y. Feng, J. Li and T. Nguyen, Application-Layer DDoS Defense with Reinforcement Learning, 2020 IEEE/ACM 28th International
Symposium on Quality of Service (IWQoS), 2020, pp. 1-10, doi : 10.1109/IWQoS49365.2020.9213026.



%############ gnn references
 
\bibitem{g_presentation} 
http://web.stanford.edu/class/cs224w/
\bibitem{GRL_Book} Citation: William L. Hamilton. (2020). Graph Representation Learning. Synthesis Lectures on Artificial Intelligence and Machine Learning, Vol. 14, No. 3 , Pages 1-159.
\bibitem{network_euclidian}
M. M. Bronstein, J. Bruna, Y. LeCun, A. Szlam and P. Vandergheynst, "Geometric Deep Learning: Going beyond Euclidean data," in IEEE Signal Processing Magazine, vol. 34, no. 4, pp. 18-42, July 2017, doi: 10.1109/MSP.2017.2693418. 

\bibitem{gnn}
Gilmer, S. Schoenholz, P. Riley, O. Vinyals, and G. Dahl. Neural message
passing for quantum chemistry. In ICML, 2017.
\bibitem{gnn_review} 
Jie Zhou, Ganqu Cui, Shengding Hu, Zhengyan Zhang, Cheng Yang, Zhiyuan Liu, Lifeng Wang, Changcheng Li, Maosong Sun,
Graph neural networks: A review of methods and applications, AI Open, Volume 1, 2020, Pages 57-81, ISSN 2666-6510, https://doi.org/10.1016/j.aiopen.2021.01.001.
\bibitem{kipf}
Kipf, Thomas N., and Max Welling. “Semi-Supervised Classification with Graph Convolutional Networks.” ArXiv:1609.02907 [Cs, Stat], Feb. 2017] .


\bibitem{GraphSAGE} 
William L. Hamilton, Rex Ying, Jure Leskovec, Inductive Representation Learning on Large Graphs, Inductive Representation Learning on Large Graphs.
\bibitem{gat} 
Veličković, P., Cucurull, G., Casanova, A., Romero, A., Liò, P. \& Bengio, Y. (2017). Graph Attention Networks. 6th International Conference on Learning Representations.

\bibitem{gnn_review1} 
Zonghan Wu, Shirui Pan, Fengwen Chen, Guodong Long, Chengqi Zhang, Philip S. Yu, A Comprehensive Survey on Graph Neural Networks.
\bibitem{gapp1}
Wu, Shiwen, Sun, Fei, Zhang, Wentao, Xie, Xu, Cui, Bin. (2022). Graph Neural Networks in Recommender Systems: A Survey. ACM Computing Surveys. 55. 10.1145/3535101.
\bibitem{jiang} 
H. Jiang, X. Xing, S. Li, J. Chen and Z. Jia, "Graph Neural Networks based Recommendation Methods in Different Scenarios: A Survey," 2022 8th Annual International Conference on Network and Information Systems for Computers (ICNISC), Hangzhou, China, 2022, pp. 660-666, doi: 10.1109/ICNISC57059.2022.00135.
\bibitem{gapp2}
Weiwei Jiang, and Jiayun Luob. Graph neural network for traffic forecasting: A survey. Expert Systems with Applications Volume, vol. 207, November 30, 2022.
\bibitem{iot}
Dong, Guimin, Mingyue Tang, Zhiyuan Wang, Jiechao Gao, Sikun Guo, Lihua Cai, Robert Gutierrez, Brad Campbell, Laura E. Barnes and Mehdi Boukhechba. “Graph Neural Networks in IoT: A Survey.” ACM Transactions on Sensor Networks 19 (2022): 1 - 50.
\bibitem{gapp3}
Zhou, Jie \& Cui, Ganqu & Zhang, Zhengyan & Yang, Cheng & Liu, Zhiyuan & Sun, Maosong. (2018). Graph Neural Networks: A Review of Methods and Applications
\bibitem{peng}
Peng, Yun & Choi, Byron & Xu, Jianliang. (2021). Graph Learning for Combinatorial Optimization: A Survey of State-of-the-Art. Data Science and Engineering. 6. 10.1007/s41019-021-00155-3.

%########## GNN-intrusion detection

\bibitem {gnnIntrisionDetection1}
Weiwei Jiang, Graph-based deep learning for communication networks: A survey, Computer Communications, Volume 185, 2022, Pages 40-54, ISSN 0140-3664, https://doi.org/10.1016/j.comcom.2021.12.015.
\bibitem {gnnIntrisionDetection2}
 Wang, Y.; Li, J.; Zhao, W.; Han, Z.; Zhao, H.; Wang, L.; He, X. N-STGAT: Spatio-Temporal Graph Neural Network Based Network Intrusion Detection for Near-Earth Remote Sensing. Remote Sens. 2023,15, 3611. https://doi.org/10.3390/rs15143611.
\bibitem {unveiling} 
D. Pujol Perich, J. R. Suarez-Varela Maci ´ a, A. Cabellos Aparicio, and ´ P. Barlet Ros, “Unveiling the potential of graph neural networks for robust intrusion detection,” in 3rd International Workshop on AI in Networks and Distributed Systems, 2021, pp. 1–7.

%################# GNN-DDoS
\bibitem{graphddos} 
Y. Li, R. Li, Z. Zhou, J. Guo, W. Yang, M. Du, and Q. Liu, “GraphDDoS: Effective DDoS Attack Detection Using Graph Neural Networks,” in 2022 IEEE 25th International Conference on Computer Supported Cooperative Work in Design (CSCWD), 2022, pp. 1275–1280.

\bibitem{Ftgnet} 
Luca Barsellotti, Lorenzo De Marinis, Filippo Cugini, Francesco Paolucci, FTG-Net: Hierarchical Flow-to-Traffic Graph Neural Network for DDoS Attack Detection, 2023 IEEE 24th International Conference on High Performance Switching and Routing (HPSR).

\bibitem{gldnet} 
W. Guo, H. Qiu, Z. Liu, J. Zhu, and Q. Wang, “GLD-Net: Deep Learning to Detect DDoS Attack via Topological and Traffic Feature Fusion,” Computational Intelligence and Neuroscience, vol. 2022.


% ######### experiments and training


\bibitem{cic} 
Univ of new Brunswick. Cic datasets https://www.unb.ca/cic/datasets/.html.
\bibitem{cross_entropy_loss} 
Anqi Mao, Mehryar Mohri, Yutao Zhong, Cross-Entropy Loss Functions: Theoretical Analysis and Applications, International Conference on Machine Learning.
\bibitem{cross_entropy_loss_1} 
 Probability for Machine Learning: Discover How To Harness Uncertainty With Python, Jason Brownlee, 2019, p. 220.
\bibitem{entropy} 
By C. E. SHANNON, A Mathematical Theory of Communication, Reprinted with corrections from The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, July, October, 1948.
\bibitem{classes_weights1} 
Prakash, Ravi \& Kumar, Karmanya. (2022). Class Weight technique for Handling Class Imbalance.
\bibitem{classes_weights2}
https://www.analyticsvidhya.com/blog/2020/10/improve-class-imbalance-class-weights/\#h-what-are-class-weights.



\end{thebibliography}

